\papertitle

### 2025-05-10 paper/main.tex
update paper

To address the need for such a middle ground, this article contributes a novel accurate, easy-to-use, and free method of minimally-congested driving travel time prediction. Our results show a substantial improvement over traditional, naïve methods of travel time prediction common in the urban planning literature. As we shall show in this article, across the Los Angeles urban area, our average trip's travel time prediction differs from Google's \enquote{gold standard} prediction by only 0.38 seconds. In contrast, simple---but common in the literature---naïve travel time prediction minimizing max-speed edge traversal times under-predicts by 183.69 seconds on average.

@@

\begin{figure*}[bt!]
    \centering
    \includegraphics[width=1.0\textwidth]{fig_workflow.jpg}
    \caption{Detailed model workflow.}\label{fig:workflow}
\end{figure*}

@@

\begin{table*}[tb!]
    \centering\small
    \caption{Out-of-sample prediction accuracy of our chosen model (random forest, bold) versus the initial naïve travel time model and the discarded alternative models, all validated against the corresponding true travel times. $n$ = 8,272 for each. The six accuracy indicators are (1) MAPE \%, (2) MAE in seconds, (3) MSE in square seconds, (4) difference-in-means ($\delta$) in seconds and its corresponding $t$-test's $p$-value, (5) APR, and (6) $R^2$. See methods for definitions.}\label{tab:validation_results}
    \tabularnums{\small
    \begin{tabular}{lrrrrrrr}
        \toprule
        Model               & MAPE  & MAE    & MSE      & $\delta$ & $p$  & APR  & $R^2$ \\
        \midrule
        Initial naïve       & 21.15 & 183.69 & 48214.06 &  -182.85 & <0.01 & 0.79 & 0.74  \\
        \textbf{Random forest}       &  \textbf{8.42} &  \textbf{75.32} & \textbf{12154.79} &     \textbf{0.38} &  \textbf{0.76} & \textbf{1.01} & \textbf{0.93}  \\
        Gradient boosting   &  7.86 &  71.99 & 12630.03 &   -19.20 & <0.01 & 0.99 & 0.93  \\
        Decision trees      &  9.00 &  80.04 & 13570.17 &     0.13 &  0.92 & 1.01 & 0.93  \\
        AdaBoost            &  8.20 &  74.04 & 12424.14 &    -9.76 & <0.01 & 1.00 & 0.93  \\
        \bottomrule
    \end{tabular}}
\end{table*}

@@

The data that support the findings of this study are available at [OFFICIAL REPOSITORY URL REDACTED FOR ANONYMOUS PEER REVIEW.\@ REVIEWERS, PLEASE USE THIS ANONYMIZED URL FOR YOUR REVIEW INSTEAD:\@ https://figshare.com/s/0acd076b1a11286ff3e4].
%https://github.com/gboeing/travel-time-prediction

>>>>

To address the need for such a middle ground, this article contributes a novel accurate, easy-to-use, and free method of minimally-congested driving travel time prediction. Our results show a substantial improvement over traditional, naïve methods of travel time prediction common in the urban planning literature. As we shall show in this article, across the Los Angeles urban area, our average trip's travel time prediction differs from Google's \enquote{gold standard} prediction by only 0.38 seconds. In contrast, simple---but common in the literature---naïve travel time prediction minimizing max-speed edge traversal times under-predicts by 183.68 seconds on average.

@@

\begin{figure*}[bt!]
    \centering
    \includegraphics[width=1.0\textwidth]{fig_workflow_larger_font.jpg}
    \caption{Detailed model workflow.}\label{fig:workflow}
\end{figure*}

@@

\begin{table*}[tb!]
    \centering\small
    \caption{Out-of-sample prediction accuracy of our chosen model (random forest, bold) versus the initial naïve travel time model and the discarded alternative models, all validated against the corresponding true travel times. $n$ = 8,272 for each. The six accuracy indicators are (1) MAPE \%, (2) MAE in seconds, (3) MSE in square seconds, (4) difference-in-means ($\delta$) in seconds and its corresponding $t$-test's $p$-value, (5) APR, and (6) $R^2$. See methods for definitions.}\label{tab:validation_results}
    \tabularnums{\small
    \begin{tabular}{lrrrrrrr}
        \toprule
        Model               & MAPE  & MAE    & MSE      & $\delta$ & $p$  & APR  & $R^2$ \\
        \midrule
        Initial naïve       & 21.15 & 183.68 & 48214.06 &  -182.85 & <0.01 & 0.79 & 0.74  \\
        \textbf{Random forest}       &  \textbf{8.41} &  \textbf{75.32} & \textbf{12154.79} &     \textbf{0.38} &  \textbf{0.76} & \textbf{1.01} & \textbf{0.93}  \\
        Gradient boosting   &  7.86 &  71.99 & 12630.03 &   -19.20 & <0.01 & 0.99 & 0.93  \\
        Decision trees      &  9.00 &  80.04 & 13570.17 &     0.13 &  0.92 & 1.01 & 0.93  \\
        AdaBoost            &  8.20 &  74.04 & 12424.14 &    -9.76 & <0.01 & 1.00 & 0.93  \\
        \bottomrule
    \end{tabular}}
\end{table*}

@@

The code and data that support the findings of this study are available at [REDACTED]. Google's terms of use forbid publicly redistributing their data: for reproducibility purposes, those training data are available directly from the authors upon request. [Reviewers: please use this anonymized URL for your review: https://figshare.com/s/0acd076b1a11286ff3e4 We have submitted the training data file separately to the journal as supplementary material for review only, to abide by the terms of use.]
%https://github.com/gboeing/travel-time-prediction
